Attrition and Achievement Gaps in Online Learning

### 2015-01-24 paper_v3.tex
updated version

<<s1coefs, echo=FALSE, fig.env='figure*', fig.pos='ht', fig.height=6, fig.width=10, fig.cap='Individual differnces in persistence and course grades showing substantial achievement gaps. An estimate of -10 indicates a 10\\% lower chance of achieving a milestone.'>>=
library(ggplot2)
res = read.csv("pred_coefs2014-1-15.csv")

res$y = factor(res$y, levels=c(">30% videos", ">50% videos", ">80% videos", ">60% pctile grade", ">30% assessments", ">50% assessments", ">80% assessments", ">80% pctile grade"))
res$y = factor(res$y, labels=c("> 30% lectures", "> 50% lectures", "> 80% lectures", "> 60th percentile grade", "> 30% assessments", "> 50% assessments", "> 80% assessments", "> 80th percentile grade"))

res$coef2 = factor(res$coef2, levels=rev(c("Intercept", "Gender:F", "Age (norm.)", "Educ.:Bachelor+", "Prior Exp.", "Intent:All", "Hours:2-4", "Hours:4+ ", "Prior MOOCs:1-3", "Prior MOOCs:4+ ", "Geo:Europe", "Geo:Oceania", "Geo:Africa", "Geo:Asia", "Geo:Latin America")))
res$coef2 = factor(res$coef2, labels=rev(c("Intercept", "Gender: Female", "Age (normalized)", "Education: Bachelor+", "Prior Experience", "Intent to complete", "Hours: 2-4", "Hours: 4+ ", "Prior MOOCs: 1-3", "Prior MOOCs: 4+ ", "Location: Europe", "Location: Oceania", "Location: Africa", "Location: Asia", "Location: Latin America")))

# This is what we submitted
# ggplot(subset(res, coef2!="Intercept"), 
#        aes(coef2, 100*est, ymin=100*(est-1.96*se), ymax=100*(est+1.96*se), color=factor(sign))) + 
#   geom_hline(xintercept=0) + geom_pointrange() + facet_wrap(~y, ncol=4) + coord_flip() + 
#   theme_minimal() + theme(legend.position="none",axis.title=element_text(vjust=0)) + 
#   scale_color_manual(values=c('red','grey30','green3')) +
#   labs(y="Predicted change in probability of achieving each `milestone` (95% CIs)", x="")

# This is what it should be
ggplot(subset(res, coef2!="Intercept"), 
  aes(coef2, 100*(exp(est)-1), ymin=100*(exp(est-1.96*se)-1), ymax=100*(exp(est+1.96*se)-1), color=factor(sign))) + 
  geom_hline(yintercept=0) + geom_pointrange() + facet_wrap(~y, ncol=4) + coord_flip() + 
  theme_minimal() + theme(legend.position="none",
                          axis.title=element_text(vjust=0),
                          panel.border=element_rect(fill="transparent",size=.5,color="grey50")) + 
  scale_color_manual(values=c('black','grey50','black')) +
  scale_y_continuous(breaks=c(-60,-40,-20,0,20,40,60,80)) +
  labs(y="Predicted change in probability of achieving each milestone (95% CIs)", x="")
coefs = subset(res, coef2=="Intercept", select=c("est","y"))  # should report these in the text of reference
coefs$exp_est = round(100*exp(coefs$est)) # estimates seem high
# 1   -1.63          > 30% lectures      20
# 16  -2.20          > 50% lectures      11
# 31  -3.13          > 80% lectures       4
# 46  -1.00       > 30% assessments      37
# 61  -1.38       > 50% assessments      25
# 76  -1.73       > 80% assessments      18
# 91  -1.21 > 60th percentile grade      30
# 106 -1.51 > 80th percentile grade      22
@

@@

<<s1reason, echo=FALSE, fig.env='figure*', fig.pos='ht', fig.height=4, fig.width=10, fig.cap='Distribution of self-reported reasons for attrition across 20 MOOCs.'>>=
library(ggplot2)
x = read.csv("drop_reasons.csv")
ggplot(x, aes(reorder(variable, value, median), round(100*value))) + 
  geom_boxplot() + coord_flip() + theme_bw() + labs(y="Proportion of learners in the course who selected the reason",x="") + scale_y_continuous(breaks=c(0,10,20,30,40,50,60,70))
@

>>>>

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure*}[ht]

\includegraphics[width=\maxwidth]{figure/s1coefs} \caption[Individual differnces in persistence and course grades showing substantial achievement gaps]{Individual differnces in persistence and course grades showing substantial achievement gaps. An estimate of -10 indicates a 10\% lower chance of achieving a milestone.\label{fig:s1coefs}}
\end{figure*}

\end{knitrout}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure*}[ht]

\includegraphics[width=\maxwidth]{figure/s1reason} \caption[Distribution of self-reported reasons for attrition across 20 MOOCs]{Distribution of self-reported reasons for attrition across 20 MOOCs.\label{fig:s1reason}}
\end{figure*}

\end{knitrout}
